---
title: "Starbucks Data on Leaflet"
author: "Arti Annaswamy"
date: "10/11/2016"
output: html_document
---

# Starbucks location visualization

### Data source: https://opendata.socrata.com/Business/All-Starbucks-Locations-in-the-US/txu4-fsic

### Load up packages

```{r, results='hide'}
library(ggplot2)
library(dplyr)
library(lubridate)
library(leaflet)
library(tm)
library(SnowballC)
library(wordcloud)
```

### Read in data

```{r}
setwd("/Users/Arti/GitHub/starbucks-leaflet")
sbux <- read.csv("All_Starbucks_Locations_in_the_US.csv", stringsAsFactors = FALSE)
```

### Feature selection and formatting

```{r}
glimpse(sbux)

# Rename fields
sbux <- sbux %>% rename(Features.Products = Features...Products) %>% 
  rename(Features.Service = Features...Service) %>% 
  rename(Features.Stations = Features...Stations)

# Create Station category and fill-in blanks
sbux$Features.Stations[sbux$Features.Stations == ""] <- "Not Drive-Through"
sbux$Features.Stations <- as.factor(sbux$Features.Stations)
table(sbux$Features.Stations)

# Create Ownership Type factor
table(sbux$Ownership.Type)
sbux$Ownership.Type <- as.factor(sbux$Ownership.Type)

```

### Franchise Type Categorizations

```{r}
# Explore Licensed locations by franchise type
sbux$License.Franchise <- NA

# University
sbux$License.Franchise[grepl("college", tolower(sbux$Name)) | grepl("university", tolower(sbux$Name)) | grepl("campus", tolower(sbux$Name)) | grepl("student", tolower(sbux$Name)) | grepl("library", tolower(sbux$Name)) | grepl("univ", tolower(sbux$Name))] <- "Education"

# Travel Centers, Airports, Conventions
sbux$License.Franchise[grepl("airport", tolower(sbux$Name)) | grepl("concourse", tolower(sbux$Name)) | grepl("terminal", tolower(sbux$Name)) | grepl("ohare", tolower(sbux$Name)) | grepl("baggage", tolower(sbux$Name)) | grepl("term ", tolower(sbux$Name)) | grepl("security ", tolower(sbux$Name))] <- "Airport"

sbux$License.Franchise[grepl("travel plaza", tolower(sbux$Name)) | grepl("travel oasis", tolower(sbux$Name)) | grepl("travel center", tolower(sbux$Name))] <- "Travel Centers"

sbux$License.Franchise[grepl("convention", tolower(sbux$Name))] <- "Convention Center"

sbux$License.Franchise[grepl("food court", tolower(sbux$Name)) | grepl("food ct", tolower(sbux$Name))] <- "Food Court"

sbux$License.Franchise[grepl("food pyramid", tolower(sbux$Name)) | grepl("food pavilion", tolower(sbux$Name)) | grepl("food store", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"

# Town Centers, Regional Centers
sbux$License.Franchise[grepl("town center", tolower(sbux$Name)) | grepl("towne center", tolower(sbux$Name)) | grepl("regional center", tolower(sbux$Name)) | grepl("trade center", tolower(sbux$Name)) | grepl("community center", tolower(sbux$Name)) | grepl("shopping center", tolower(sbux$Name)) | grepl(" center", tolower(sbux$Name)) | grepl(" ctr", tolower(sbux$Name))] <- "Shopping - Misc"
sbux$License.Franchise[grepl(" city", tolower(sbux$Name)) | grepl(" lobby", tolower(sbux$Name)) | grepl(" courtyard", tolower(sbux$Name)) | grepl(" mall", tolower(sbux$Name))] <- "Shopping - Misc"

# General categories
sbux$License.Franchise[grepl("hospital", tolower(sbux$Name)) | grepl("medical", tolower(sbux$Name))] <- "Hospital"
sbux$License.Franchise[grepl("beach", tolower(sbux$Name))] <- "Beach"
sbux$License.Franchise[grepl("casino", tolower(sbux$Name)) | grepl("harrah", tolower(sbux$Name))] <- "Hotel"
sbux$License.Franchise[grepl("fort ", tolower(sbux$Name)) | grepl("ft. ", tolower(sbux$Name))] <- "Army Bases"

# Specific Locations
sbux$License.Franchise[grepl("safeway", tolower(tolower(sbux$Name)))] <- "Supermarket/Grocery - Safeway"
sbux$License.Franchise[grepl("target", tolower(sbux$Name))] <- "Supermarket/Grocery - Target"
sbux$License.Franchise[grepl("albertson", tolower(sbux$Name))] <- "Supermarket/Grocery - Albertsons"
sbux$License.Franchise[grepl("vons", tolower(sbux$Name)) | grepl("von's", tolower(sbux$Name))] <- "Supermarket/Grocery - Vons"
sbux$License.Franchise[grepl("fred meyer", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("kroger", tolower(sbux$Name))] <- "Supermarket/Grocery - Kroger"
sbux$License.Franchise[grepl("harris teeter", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("king sooper", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("dominick", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("tom thumb", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("stop", tolower(sbux$Name)) & grepl("shop", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("hy", tolower(sbux$Name)) & grepl("vee", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("farm fresh", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("piggly wiggly", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("qfc", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("bashas", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("city market", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("dillon", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("genuardi", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("ingles", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("martin's", tolower(sbux$Name)) | grepl("martins", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("meijer", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("price cutter", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("randall", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("smith's", tolower(sbux$Name)) | grepl("smiths", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("spartan", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("a&p", tolower(sbux$Name)) | grepl("a & p", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"
sbux$License.Franchise[grepl("giant ", tolower(sbux$Name)) | grepl("giant-", tolower(sbux$Name))] <- "Supermarket/Grocery - Misc"

# Shopping
sbux$License.Franchise[grepl("macy", tolower(sbux$Name))] <- "Shopping - Macy's"
sbux$License.Franchise[grepl("barnes", tolower(sbux$Name))] <- "Shopping - Barnes & Noble"
sbux$License.Franchise[grepl("fry", tolower(sbux$Name))] <- "Shopping - Fry's"

# Hotels
sbux$License.Franchise[grepl("marriott", tolower(sbux$Name)) | grepl("hilton", tolower(sbux$Name)) | grepl("hotel", tolower(sbux$Name)) | grepl("sheraton", tolower(sbux$Name)) | grepl("westin", tolower(sbux$Name)) | grepl("hyatt", tolower(sbux$Name))] <- "Hotel"

sbux$License.Franchise[sbux$Ownership.Type == "Company Owned"] <- "Starbucks-Owned"

sbux$License.Franchise[is.na(sbux$License.Franchise) == TRUE] <- "TBD"

# For quick checks

# sbux$Name[grepl("qfc", tolower(sbux$Name))]
# table(sbux$License.Franchise)

# Franchises table

franchises <- sbux %>% 
  group_by(License.Franchise) %>% 
  summarize(Count = n()) %>% 
  arrange(desc(Count))

```

### Wordcloud 

```{r}

# Create wordcloud of Starbucks Licensed Store names
sbuxCorpus <- Corpus(VectorSource(sbux$Name))
sbuxCorpus <- tm_map(sbuxCorpus, PlainTextDocument)
sbuxCorpus <- tm_map(sbuxCorpus, removePunctuation)
sbuxCorpus <- tm_map(sbuxCorpus, removeWords, stopwords('english'))
wordcloud(sbuxCorpus, max.words = 100, random.order = FALSE)

```

### Mapping with Leaflet

#### Cluster-type map by Drive-Through / Not Drive-Through
```{r}
sbux_cluster <- leaflet() %>% 
  addTiles() %>% 
  setView(-101.140, 40.497, zoom = 5) %>%
  addCircleMarkers(data = sbux, lng = ~ Longitude, lat = ~ Latitude, 
                   radius = 7, 
                   color = ~ ifelse(Features.Stations == 'Drive-Through', 
                                    'seagreen', 'purple'),
                   popup = ~ paste(Name, " - ", Ownership.Type, " (", Features.Stations, ")", sep = ""), 
                   clusterOptions = markerClusterOptions())
sbux_cluster
```

#### Individual locations by Drive-Through / Not Drive-Through
```{r}

pal_station <- colorFactor(c("seagreen", "goldenrod"), domain = c("Drive-Through", "Not Drive-Through"))

sbux_station <- leaflet() %>% 
  addTiles() %>% 
  setView(-101.140, 40.497, zoom = 5) %>%
  addCircleMarkers(data = sbux, lng = ~ Longitude, lat = ~ Latitude, 
                   radius = 5, 
                   color = ~ ifelse(Features.Stations == 'Drive-Through', 
                                    'seagreen', 'goldenrod'), popup = ~ paste(Name, " - ", Ownership.Type, " (", Features.Stations, ")", sep = "")) %>% addLegend("bottomright", pal = pal_station, values = sbux$Features.Stations, title = "Starbucks Station Type", opacity = 1)

sbux_station
```

#### Individual locations by Starbucks Owned vs Licensed
```{r}

pal_ownership <- colorFactor(c("seagreen", "coral"), domain = c("Company Owned", "Licensed"))

sbux_ownership <- leaflet() %>% 
  addTiles() %>% 
  setView(-101.140, 40.497, zoom = 5) %>%
  addCircleMarkers(data = sbux, lng = ~ Longitude, lat = ~ Latitude, 
                   radius = 5, 
                   color = ~ ifelse(Ownership.Type == 'Company Owned', 
                                    'seagreen', 'coral'), popup = ~ paste(Name, "  (", Ownership.Type, ")", sep = "")) %>% addLegend("bottomright", pal = pal_ownership, values = sbux$Ownership.Type, title = "Starbucks Ownership Type", opacity = 1)
sbux_ownership
```

#### Individual locations by type of Licensed
```{r}

sbux_franchise <- leaflet() %>% 
  addTiles() %>% 
  setView(-101.140, 40.497, zoom = 5) %>%
  addCircleMarkers(data = subset(sbux, sbux$License.Franchise == "Starbucks-Owned"), group = "Starbucks-Owned", lng = ~ Longitude, lat = ~ Latitude, radius = 7, fillOpacity = 0.5, color = "seagreen", popup = ~ paste(Name, "  (", License.Franchise, ")", sep = "")) %>% 
  addCircleMarkers(data = subset(sbux, sbux$License.Franchise == "Supermarket/Grocery - Target"), group = "Supermarket/Grocery - Target", lng = ~ Longitude, lat = ~ Latitude, radius = 7, fillOpacity = 0.5, color = "red", popup = ~ paste(Name, "  (", License.Franchise, ")", sep = "")) %>% 
  addCircleMarkers(data = subset(sbux, sbux$License.Franchise == "Supermarket/Grocery - Safeway"), group = "Supermarket/Grocery - Safeway", lng = ~ Longitude, lat = ~ Latitude, radius = 7, fillOpacity = 0.5, color = "blue", popup = ~ paste(Name, "  (", License.Franchise, ")", sep = "")) %>% 
  addCircleMarkers(data = subset(sbux, sbux$License.Franchise == "Education"), group = "Education", lng = ~ Longitude, lat = ~ Latitude, radius = 7, fillOpacity = 0.5, color = "purple", popup = ~ paste(Name, "  (", License.Franchise, ")", sep = "")) %>% 
  addCircleMarkers(data = subset(sbux, sbux$License.Franchise == "Supermarket/Grocery - Kroger"), group = "Supermarket/Grocery - Kroger", lng = ~ Longitude, lat = ~ Latitude, radius = 7, fillOpacity = 0.5, color = "deepskyblue", popup = ~ paste(Name, "  (", License.Franchise, ")", sep = "")) %>% 
  addCircleMarkers(data = subset(sbux, sbux$License.Franchise == "Supermarket/Grocery - Albertsons"), group = "Supermarket/Grocery - Albertsons", lng = ~ Longitude, lat = ~ Latitude, radius = 7, fillOpacity = 0.5, color = "skyblue", popup = ~ paste(Name, "  (", License.Franchise, ")", sep = "")) %>% 
  addCircleMarkers(data = subset(sbux, sbux$License.Franchise == "Supermarket/Grocery - Misc"), group = "Supermarket/Grocery - Misc", lng = ~ Longitude, lat = ~ Latitude, radius = 7, fillOpacity = 0.5, color = "slateblue", popup = ~ paste(Name, "  (", License.Franchise, ")", sep = "")) %>% 
  addCircleMarkers(data = subset(sbux, sbux$License.Franchise == "Supermarket/Grocery - Vons"), group = "Supermarket/Grocery - Vons", lng = ~ Longitude, lat = ~ Latitude, radius = 7, fillOpacity = 0.5, color = "royalblue", popup = ~ paste(Name, "  (", License.Franchise, ")", sep = "")) %>% 
  addCircleMarkers(data = subset(sbux, sbux$License.Franchise == "Airport"), group = "Airport", lng = ~ Longitude, lat = ~ Latitude, radius = 7, fillOpacity = 0.5, color = "chartreuse", popup = ~ paste(Name, "  (", License.Franchise, ")", sep = "")) %>% 
  addCircleMarkers(data = subset(sbux, sbux$License.Franchise == "Hotel"), group = "Hotel", lng = ~ Longitude, lat = ~ Latitude, radius = 7, fillOpacity = 0.5, color = "salmon", popup = ~ paste(Name, "  (", License.Franchise, ")", sep = "")) %>% 
  addCircleMarkers(data = subset(sbux, sbux$License.Franchise == "Shopping - Misc"), group = "Shopping - Misc", lng = ~ Longitude, lat = ~ Latitude, radius = 7, fillOpacity = 0.5, color = "orange", popup = ~ paste(Name, "  (", License.Franchise, ")", sep = "")) %>% 
  addLayersControl(
    overlayGroups = c("Starbucks-Owned", "Supermarket/Grocery - Target", "Supermarket/Grocery - Safeway", "Supermarket/Grocery - Misc", "Education", "Supermarket/Grocery - Kroger", "Supermarket/Grocery - Albertsons", "Supermarket/Grocery - Vons", "Airport", "Hotel", "Shopping - Misc"),
    options = layersControlOptions(collapsed = FALSE)
  )

sbux_franchise

```